---
output: 
  html_document:
    keep_md: true
---

Title: "Reproducible Research: Peer Assessment 1"
=======================================================
```{r setoptions,echo=FALSE,results = "hide"}
library(knitr)
library(dplyr)
opts_chunk$set(echo=FALSE, results = "hide")
```

## Loading and preprocessing the data
```{r loadAndPreProcessData,echo=TRUE,results='hide'}
        filePath="C:\\_R\\4 -Reproducible Research\\Assignments\\RepData_PeerAssessment1\\activity.csv"
        
                
                #check if file exists
                if(!file.exists(filePath))
                {
                       print(sprintf("The file %s does not exist. Cannot proceed with data load.",filePath))         
                }
                #load the data
                origdata <- tbl_df(read.csv(filePath))
                data <- filter(origdata,!is.na(steps))
        
```


## What is mean total number of steps taken per day?
#### Histogram of Total Steps
```{r step1hist,echo=TRUE,results='asis'}
activity <- group_by(data,date)
#Calculate the Total,Mean and Median of the Steps taken per day
summary <- summarize(activity,Total=as.numeric(sum(steps)),Mean=mean(steps),Median=median(steps))
hist(summary$Total, breaks = 12, col = "lightblue", border = "blue",xlab="Total Steps",main="Fig 1: Histogram of Total Steps")
```

#### Table of Mean and Median

```{r step1TableMean,echo=TRUE,results='asis'}
library(xtable)
tab <- xtable(select(summary,date,Mean,Median))
print(tab,type="html")
```

## What is the average daily activity pattern?
#### Time series plot 
```{r step2TimeSeriesPlot,echo=TRUE,results='asis'}
activity <- group_by(data,interval)
#Calculate the Total,Mean and Median of the Steps taken by interval averaged across all days
summary <- summarize(activity,Total=as.numeric(sum(steps)),Mean=mean(steps),Median=median(steps))
plot(summary$interval,summary$Mean,xlab="Interval",ylab="Average Steps",type="l",main="Fig 2: Number of Steps taken averaged across all Days.")

```

#### Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r step2WhichIntervalHasMaxSteps, echo=TRUE, results='asis'}
m <- filter(summary,Mean==max(summary$Mean))
sprintf('Interval = %i contains the maximum average Steps taken across all days.',m$interval) 
```

## Imputing missing values
#### Calculate and Report total number of missing values.
```{r step3ReportTotalMissingValues,echo=TRUE,results='asis'}
sprintf("There are %i total missing values in the original activity dataset",count(filter(origdata,is.na(steps)))$n )

```

#### Devise a strategy for filling in all of the missing values in the dataset
```{r step3FillMissingValues, echo=TRUE, results='asis'}
imputeddata<- origdata %>%
    group_by(interval)  %>%
    mutate(imputedsteps= ifelse(is.na(steps), mean(steps, na.rm=TRUE), steps))
imputeddata<-group_by(imputeddata,date)
summary <- summarize(imputeddata,Total=as.numeric(sum(imputedsteps)),Mean=mean(imputedsteps),Median=median(imputedsteps))
hist(summary$Total, breaks = 12, col = "lightblue", border = "blue",xlab="Total Steps",main="Fig 3: Histogram of Total Steps from imputed values")
```

#### Table of Mean and Median using the Imputed values

```{r step3TableMeanForImputedValues,echo=TRUE,results='asis'}
library(xtable)
tab <- xtable(select(summary,date,Mean,Median))
print(tab,type="html")
```

## Are there differences in activity patterns between weekdays and weekends?
```{r step4CreateFactorVar,echo=TRUE,results='asis'}
imputeddata <- mutate(imputeddata,day=weekdays(as.Date(date)),weekday=ifelse(day %in% c("Monday","Tuesday","Wednesday","Thursday","Friday"),"weekday","weekend"))
imputeddata <- group_by(imputeddata,weekday,interval)
summary <- summarize(imputeddata,totalsteps = sum(imputedsteps),Mean = mean(imputedsteps), Median = median(imputedsteps))

library(lattice)
xyplot(Mean ~ interval|weekday, data=summary, type='l', layout=(c(1,2)),
   main="Figure 4: Average Daily Activity Pattern by type of Day",
   ylab="Average Number of Steps Taken per Interval", xlab="5-minute Time Interval")  

```